Scale Space Edge Detection

### 2018-03-07 paper.tex
HERES UIOTEWHPIUHTEPIWU

\begin{itemize}
\item large discontinuities in luminance values
\item discontinuities in different object orientations
\item large discontinuities in the intensity gradient
\end{itemize}
\indent Some of the common methods for edge detection include the Sobel, Canny, Prewitt, Roberts, and Fuzzy Logic algorithms.
However, these methods do not yield a lot of imformation with regards to the strength of the edges detected.
The majority of these algorithms will usually convolve the image with a static matrix to caclulate edges and does adapt enough to the image.

This is why for our edge detection method, we will be using the scale space approach.
The benefit of using a scale space appraoch for edge detection, is we have the ability to classify the strength of the edges in the image.
This allows for a range of edges from very large immediate changes in intensity to very gradual.

\section{Scale Space and Its Derivatives}
To understand how exactly we detect images in the scale space, we must first define what the scale space is.
If we have a continuous function of multiple variables such as $f(x,y)$, then we define the scale space representation of such a function as 
\begin{equation}
  L(x;t) = g(x,y;t) * f(x,y)
\end{equation}
Here $t$ represents the scale parameter, and can be thought of how much smoothing is applied to the function. The function $g$ is the Gaussian kernel given by
\begin{equation}
  g(x,y;t) = \frac{1}{2 \pi t}e^{-(x^2+y^2)/(2t)}
\end{equation}
With the scale space representation defined, we can now take derivatives of it as it is a continuous well-defined function.
Spatial derivatives are relatively simple being defined as the following
\begin{equation}
  L_{x^{\alpha}y^{\beta}}(\cdot;t) = \partial_{x^{\alpha}y^{\beta}}L(\cdot;t) = g_{x^{\alpha}y^{\beta}}(\cdot;t) * f(\cdot)
\end{equation}
However, when taking the partial derivative with respect to the scale $t$, it becomes more interesting.
The scale space representation collection is a solution for the diffusion equation. Therefore is has the useful property of 
\begin{equation}
  \partial_t L = \frac{1}{2} \nabla^2 L = \frac{1}{2} (\partial_{xx} + \partial_{yy})L
\end{equation}
with the inital condition of $L(x,y;0) = f(x,y)$.
So now, scale derivatives can be represented as spatial derivatives.

@@

# Third order derivatives
Lxxx = convolve_scale_space(Lxx, Dx)
Lxxy = convolve_scale_space(Lxx, Dy)
Lxyy = convolve_scale_space(Lxy, Dy)
Lyyy = convolve_scale_space(Lyy, Dy)
\end{lstlisting}
By convolving our ando5 derivative matrix with the scale space representation, we can take spatial derivatives in both the $x$ and $y$ dimension. To give an idea of what the images look like, we have the following spatial derivatives of the image below.
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{{{Images/lena/lena_Lv_maxima_14.439}}}
  \caption{Derivative of image in $x$ direction}
  \label{lx}
\end{figure}
\begin{figure}[H]
  \centering
  \includegraphics[scale=0.5]{Images/lena/lena_final.png}
  \caption{Derivative of image in $y$ direction}
  \label{ly}
\end{figure}
Now that we have computed the spatial derivatives of the scale space representation of the image, we can check if we fulfill the conditions in (\ref{eq:c1}). 
The next step is to then calculate the edge strength derivatives neccessary to fulfill the conditions of (\ref{eq:c2}).
The code block below shows the computations for the edge strength based on the gradient magnitude and its derivatives.
\begin{lstlisting}
# Shape the scales vector to be a vector with depth
scales3 = reshape(scales, 1, 1, length(scales))

@@

\begin{table}[t]
  \caption{Gradient magnitude}
  \centering
  \tabulinesep=0.1em
  \begin{tabu}to \textwidth{cX[c,m]X[c,m]X[c,m]}
    \toprule
    & Lena & Mandrill & Peppers\\
    \midrule
    $t = 0.1$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_0.1}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_0.1}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_0.1}}}\\
    $t = 1.7$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_1.7}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_1.7}}} & \includegraphics[height=11em]{{{Images/peppers/peppers_GL_1.7}}}\\
    $t = 18.7$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_18.7}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_18.7}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_18.7}}}\\
    $t = 93.6$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_93.6}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_93.6}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_93.6}}}\\
    $t = 256.0$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_256.0}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_256.0}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_256.0}}}\\
    \bottomrule
\end{tabu}
\end{table}

>>>>

\begin{itemize}
\item large discontinuities in luminance values
\item discontinuities in different object orientations
\item large discontinuities in the intensity gradient
\end{itemize}
\indent Some of the common methods for edge detection include the Sobel, Canny, Prewitt, Roberts, and Fuzzy Logic algorithms.
However, these methods do not yield a lot of information with regards to the strength of the edges detected.
The majority of these algorithms will usually convolve the image with a static matrix to calculate edges and does adapt enough to the image.

This is why for our edge detection method, we will be using the scale space approach.
The benefit of using a scale space approach for edge detection, is we have the ability to classify the strength of the edges in the image.
This allows for a range of edges from very large immediate changes in intensity to very gradual.

\section{Scale Space and Its Derivatives}
To understand how exactly we detect images in the scale space, we must first define what the scale space is.
If we have a continuous function of multiple variables such as $f(x,y)$, then we define the scale space representation of such a function as 
\begin{equation}
  L(x;t) = g(x,y;t) * f(x,y)
\end{equation}
Here $t$ represents the scale parameter, and can be thought of how much smoothing is applied to the function. The function $g$ is the Gaussian kernel given by
\begin{equation}
  g(x,y;t) = \frac{1}{2 \pi t}e^{-(x^2+y^2)/(2t)}
\end{equation}
With the scale space representation defined, we can now take derivatives of it as it is a continuous well-defined function.
Spatial derivatives are relatively simple being defined as the following
\begin{equation}
  L_{x^{\alpha}y^{\beta}}(\cdot;t) = \partial_{x^{\alpha}y^{\beta}}L(\cdot;t) = g_{x^{\alpha}y^{\beta}}(\cdot;t) * f(\cdot)
\end{equation}
However, when taking the partial derivative with respect to the scale $t$, it becomes more interesting.
The scale space representation collection is a solution for the diffusion equation. Therefore is has the useful property of 
\begin{equation}
  \partial_t L = \frac{1}{2} \nabla^2 L = \frac{1}{2} (\partial_{xx} + \partial_{yy})L
\end{equation}
with the initial condition of $L(x,y;0) = f(x,y)$.
So now, scale derivatives can be represented as spatial derivatives.

@@

# Third order derivatives
Lxxx = convolve_scale_space(Lxx, Dx)
Lxxy = convolve_scale_space(Lxx, Dy)
Lxyy = convolve_scale_space(Lxy, Dy)
Lyyy = convolve_scale_space(Lyy, Dy)
\end{lstlisting}
By convolving our ando5 derivative matrix with the scale space representation, we can take spatial derivatives in both the $x$ and $y$ dimension. To give an idea of what the images look like, we have the following spatial derivatives of the image below.
% \begin{figure}[H]
%   \centering
%   \includegraphics[scale=0.5]{{{Images/lena/lena_Lv_maxima_14.439}}}
%   \caption{Derivative of image in $x$ direction}
%   \label{lx}
% \end{figure}
% \begin{figure}[H]
%   \centering
%   \includegraphics[scale=0.5]{Images/lena/lena_final.png}
%   \caption{Derivative of image in $y$ direction}
%   \label{ly}
% \end{figure}
Now that we have computed the spatial derivatives of the scale space representation of the image, we can check if we fulfill the conditions in (\ref{eq:c1}). 
The next step is to then calculate the edge strength derivatives neccessary to fulfill the conditions of (\ref{eq:c2}).
The code block below shows the computations for the edge strength based on the gradient magnitude and its derivatives.
\begin{lstlisting}
# Shape the scales vector to be a vector with depth
scales3 = reshape(scales, 1, 1, length(scales))

@@

\begin{table}[H]
  \caption{Gradient magnitude}
  \centering
  \tabulinesep=0.1em
  \begin{tabu}to \textwidth{cX[c,m]X[c,m]X[c,m]}
    \toprule
    & Lena & Mandrill & Peppers\\
    \midrule
    $t = 0.1$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_0.1}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_0.1}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_0.1}}}\\
    $t = 1.7$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_1.7}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_1.7}}} & \includegraphics[height=11em]{{{Images/peppers/peppers_GL_1.7}}}\\
    $t = 18.7$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_18.7}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_18.7}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_18.7}}}\\
    $t = 93.6$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_93.6}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_93.6}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_93.6}}}\\
    $t = 256.0$ & \includegraphics[height=11em]{{{Images/lena/lena_GL_256.0}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GL_256.0}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GL_256.0}}}\\
    \bottomrule
\end{tabu}
\end{table}
\begin{table}[H]
  \caption{First scale derivative of gradient magnitude}
  \centering
  \tabulinesep=0.1em
  \begin{tabu}to \textwidth{cX[c,m]X[c,m]X[c,m]}
    \toprule
    & Lena & Mandrill & Peppers\\
    \midrule
    $t = 0.1$ & \includegraphics[height=11em]{{{Images/lena/lena_GLt_0.1}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GLt_0.1}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GLt_0.1}}}\\
    $t = 1.7$ & \includegraphics[height=11em]{{{Images/lena/lena_GLt_1.7}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GLt_1.7}}} & \includegraphics[height=11em]{{{Images/peppers/peppers_GLt_1.7}}}\\
    $t = 18.7$ & \includegraphics[height=11em]{{{Images/lena/lena_GLt_18.7}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GLt_18.7}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GLt_18.7}}}\\
    $t = 93.6$ & \includegraphics[height=11em]{{{Images/lena/lena_GLt_93.6}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GLt_93.6}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GLt_93.6}}}\\
    $t = 256.0$ & \includegraphics[height=11em]{{{Images/lena/lena_GLt_256.0}}} & \includegraphics[height=11em]{{{Images/mandrill/mandrill_GLt_256.0}}} &  \includegraphics[height=11em]{{{Images/peppers/peppers_GLt_256.0}}}\\
    \bottomrule
\end{tabu}
\end{table}
